---
title: "US Population Trends"
author: "Radan Ganchev"
date: "8/20/2017"
output: ioslides_presentation
---

## State Population Data

 - The **US Census Bureau** provides variuos census data and population estimates.
 - Population estimates for the years 2010-2016 are available at this site: https://www.census.gov/data/tables/2016/demo/popest/nation-total.html
 - Using this data we can estimate the current population trends in the US states.
 - Quick look at the data
```{r}
population <- read.csv("data/population.csv")
names(population)
```

## Sample Population Trend

 - Using this data, we can calculate the population trend in any given state by fitting a linear regression line
```{r, echo = FALSE}
years <- 2010:2016
state <- population[population$State == "Idaho",]
yearlyPop <- c(state$X2010, state$X2011, state$X2012, state$X2013, state$X2014, state$X2015, state$X2016)
plot(years, yearlyPop, type="l", col="red", lwd=2, yaxt="n",
                     main = paste("Population Trend for", state$State),
                     xlab = "Year",
                     ylab = "Population")
                legend("bottomright",
                       legend = c("Population", "Regression Line"),
                       lty = c(1, 1),
                       lwd = c(2, 2),
                       col = c("red", "blue"))
                axis(2, at=axTicks(2), labels=sprintf("%.2fM", axTicks(2) / 1000000))
                abline(lm(yearlyPop ~ years), col="blue", lwd=2)
```

## Shiny Web App

 - The Shiny web application deployed [here](https://rganchev.shinyapps.io/us_population/) allows users to easily see the population trends in all american states.
 
 - The interactive map can be used to access all relevant data and display a graphical representation of the trends.

```{r fig.height=3.5, echo=FALSE}
library(png)
library(grid)
img <- readPNG("webapp.png")
grid.raster(img)
```

# Thanks!

